### 2013-09-28 math_review.tex

\subsection{Coordinate Systems\label{sec:notation}}

\subsubsection{Time Derivatives}

\begin{align*}
\dot{a}& = \frac{da}{dt}\\
%
\ddot{a}& = \frac{d^2a}{dt^2}
\end{align*}

\subsubsection{Radius of Curvature}

The radius of curvature, $R$, of a curve at a point is a measure of the radius of the circular arc which best approximates the curve at that point.

\begin{align}
R& = \abs{\frac{\paren{\dot{x}^2 + \dot{y}^2}^{3/2}}{\dot{x}\ddot{y} - \dot{y}{\ddot{x}}}}\\
%
&= \abs{\frac{s^2}{a_t}}\tag{for $a_c = $}
\end{align}

\subsection{Scalar Calculus in Polar Coordinates}

\begin{align}
s& = \frac{dl}{dt}\\
%
a_t& = \frac{ds}{dt}\\
\end{align}

\subsection{Vector Calculus in Polar Coordinates}

\begin{align}
\boldsymbol{x}& = r \boldsymbol{\hat{r}}\\
%
\boldsymbol{\vec{v}}& = \dot{r} \boldsymbol{\hat{r}} + r \dot{\phi} \boldsymbol{\hat{\phi}}\\
%
\label{eq:a_vec_def}\boldsymbol{\vec{a}}& = \paren{\ddot{r} - r \dot{\phi}^2} \boldsymbol{\hat{r}} + \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}} \boldsymbol{\hat{\phi}}
\end{align}

\subsection{Relations}

\begin{align}
l& = \int_{t=0}^{T} \norm{\boldsymbol{\vec{v}} \, } \; dt\\
%
s& = \norm{\boldsymbol{\vec{v}}}\\
%
a_t& = \norm{\boldsymbol{\vec{a}}} \cdot \boldsymbol{\hat{v}} = \norm{\boldsymbol{\vec{a}}} \cdot \frac{\boldsymbol{\vec{v}}}{\norm{\boldsymbol{\vec{v}}}}\\
%
a_c& = \norm{\boldsymbol{\vec{a}}} \times \boldsymbol{\hat{v}} = \norm{\boldsymbol{\vec{a}}} \times \frac{\boldsymbol{\vec{v}}}{\norm{\boldsymbol{\vec{v}}}}
\end{align}