### 2013-09-28 math_review.tex
Add more proofs.

\subsection{Coordinate Systems}

\subsubsection{Radius of Curvature}

The radius of curvature, $R$, of a curve at a point is a measure of the radius of the circular arc which best approximates the curve at that point.

For $a_t = 0$

\[
R= \abs{\frac{v^2}{a_c}}
\]

\subsection{Vector Calculus in Polar Coordinates}

\begin{align*}
\boldsymbol{x}& = r \boldsymbol{\hat{r}}\\
\boldsymbol{\vec{v}}& = \dot{r} \boldsymbol{\hat{r}} + r \dot{\phi} \boldsymbol{\hat{\phi}}\\
\label{eq:a_vec_def}\boldsymbol{\vec{a}}& = \paren{\ddot{r} - r \dot{\phi}^2} \boldsymbol{\hat{r}} + \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}} \boldsymbol{\hat{\phi}}
\end{align*}