### 2013-09-29 cone_theorems.tex
Add cone proofs.

Now we are finally able to tackle the model problem of optimal paths around cones. The cone problem we will solve will involve only 3 cones, but first we will solve a simpler but related problem of minimizing the path length around a quadrilateral.

\subsection{Quadrilateral}

We would like to minimize the path length around the quadrilateral shown in Figure \ref{fig:quad}.

\includefig{quad.eps}{Quadrilateral path. A $\to$ B $\to$ C $\to$ D}{quad}

We will constrain the problem such that it is similar to the problem of a particle traveling around a cone. Sides $\overline{BC}$, $\overline{DA}$, and $\overline{EF}$ have fixed lengths $l_1$, $l_2$, and $w$ respectively. Points E and F bisect $\overline{DA}$ and $\overline{BC}$ respectively. The goal is

\[
\argmin_{\theta} \bracket{l_{AB} + l_{BC} + l_{CD}}
\]

\begin{align}
  x_c& = w - \frac{l2}{2 \cos{\theta}}\\
  y_c& = \frac{l_2}{2} - \frac{l_2}{2 \sin{\theta}}\\
  l_{AB}& = \sqrt{x_c^2 + y_c^2}\\
  & = \sqrt{ \paren{\frac{l_1}{2} - \frac{l_2}{2 \cos\theta}}^2 + \paren{w - \frac{l_2}{2 \sin\theta}}^2 }\\
  l_{BC}& = l2\\
  l_{CD}& = \sqrt{\paren{\frac{l_1}{2} + \frac{l_2}{2 \cos\theta}}^2 + \paren{w - \frac{l_2}{2 \sin\theta}}^2 }
\end{align}

Since we know that $l_{BC}$ is fixed the problem reduces to

\[
\argmin_{\theta} \bracket{l_{AB} + l_{BC} + l_{CD}}
\]

Because all distances are positive, we can use Lemma \ref{lemma:conv-argmin}, defined in the Appendix, to simplify our problem:

If we define $l_{AB} = \sqrt{z_{AB}}$ and $l_{CD} = \sqrt{z_{CD}}$, we can use \ref{lemma:conv-argmin} with $k(t) = \sqrt{t}$ to simplify our problem to

\begin{multline}
  \argmin_{\theta} \left[ 2\paren{\frac{l_1}{2} - \frac{l_2}{2 \sin\theta}}^2 \right.\\
  \left. + \paren{w - \frac{l_2}{2 \cos\theta}}^2 + \paren{w + \frac{l_2}{2 \cos\theta}}^2 \right]
\end{multline}

Now, we can expand out our expression and use Lemma the fact that $\sin^2 \theta + \cos^2 \theta = 1$ to obtain a much simpler (but equivalent) minimization problem:
\begin{eqnarray}
  \argmin_{\theta} \frac{l_1^2}{2} + \frac{l_2^2}{2} + 2d^2 - l_1 l_2 \sin \theta
\end{eqnarray}

We note that $l_1, l_2,$ and $d$ are all constants which are given to us in the problem. Therefore, the minimization problem really boils down to
\begin{eqnarray}
  \argmin_{\theta} - l_1 l_2 \sin \theta = \argmax_{\theta} \sin \theta
\end{eqnarray}

% -----------------------------------------------------------------------------

% 3 cones, constant speed
\begin{theorem}

Given a particle with zero tangential acceleration and bounded centripetal acceleration, $a_c \le a_{c,max}$, and a cone setup consisting of 3 cones, A, B, and C, at locations $\bvec{X_A}$, $\bvec{X_B}$, and $\bvec{X_C}$ as is shown in Figure \ref{fig:3-cones}.

\includefig{3_cones.eps}{3 Cone setup}{3-cones}

The optimal path [TODO]

\end{theorem}

\begin{proof}
Since speed is constant, the optimal path will be the shortest path. The optimal path is of the form, tr. Thus, the problem reduces to a trigonometric problem.

\includefig{cones_fix_s.eps}{Position of center of turn}{cones-fix-s}

The turn must touch the cone B, otherwise the trajectory will either not go around the cone, or be longer than necessary. Therefor, the center of the turn must be a distance $r$ away from

\includefig{cones_fix_s_2.eps}{Path length}{cones-fix-s-2}

\begin{align*}
x_c = x_B - r \cos{\alpha_c}\\
y_c = y_B - r \sin{\alpha_c}\\
l_1& = \sqrt{d_c - r}\\
l_2& = r \paren{\alpha_c - \tan^{-1}\paren{\frac{y_c}{x_c}}}\\
\end{align*}
\end{proof}