### 2013-09-29 introduction.tex

Our model problem is a particle moving in $\R^2$ with constrained acceleration. The particle must navigate around cones (points in $\R^2$) to reach a final position. The goal is to minimize the total time spend navigating to the final position.

In this paper, we first present a polar coordinate representation of a particle's position around a point and derive a set of differential equations governing the motion of the particle in this coordinate system. This allows us to devise some simple lemmas about the motion of the particle.

After developing an intuition for particles and particle paths, we then move on to the more complicated problem of deriving governing rules for optimal paths between points.

Finally, we are able to tackle the problem of finding an optimal path around cones.