### 2013-09-29 paper.tex

\documentclass[12pt]{amsart}   % LaTeX with AMS style; 12 point for old eyes

\usepackage{amsmath,amssymb,amsfonts}   % For better support of math
\usepackage{graphicx}	        % Enable for eps and pdf figures, if they occur
\usepackage{caption}
\usepackage{float}
\usepackage{subcaption}
\usepackage{hyperref} % Enable embedded hyperlinks.
\hypersetup{
	hidelinks, colorlinks, linkcolor=black, citecolor=black, urlcolor=red
}

% Commands to force sequential numbering:

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{examples}[theorem]{Examples}
\newtheorem{remarks}[theorem]{Remarks}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}

% Define abs norm, paren, bracket, cbracket, and innerproduct
\usepackage{mathtools}
\DeclarePairedDelimiter\tempabs{\lvert}{\rvert}
\DeclarePairedDelimiter\tempnorm{\lVert}{\rVert}
\DeclarePairedDelimiter\tempinnerproduct{\langle}{\rangle}
\DeclarePairedDelimiter\tempparen{(}{)}
\DeclarePairedDelimiter\tempbracket{[}{]}
\DeclarePairedDelimiter\tempcbracket{\{}{\}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
% Swap * functionality
\makeatletter
\def\abs{\@ifstar{\tempabs}{\tempabs*}}
\def\norm{\@ifstar{\tempnorm}{\tempnorm*}}
\def\innerproduct{\@ifstar{\tempinnerproduct}{\tempinnerproduct*}}
\def\paren{\@ifstar{\tempparen}{\tempparen*}}
\def\bracket{\@ifstar{\tempbracket}{\tempbracket*}}
\def\cbracket{\@ifstar{\tempcbracket}{\tempcbracket*}}
\makeatother

% blackboard bold , for ``complex,'' etc
\newcommand{\C}{\mathbb C} 
\newcommand{\R}{\mathbb R} 
\newcommand{\Z}{\mathbb Z}
\newcommand{\Q}{\mathbb Q}
\newcommand{\N}{\mathbb N}

% Add fig with caption command. Inputs: {path}{caption}{label}
\newcommand{\includefig}[3]{
  \begin{figure}[H]
    \begin{center}
      \includegraphics[width=4in,height=3in,keepaspectratio]{#1}
    \end{center}
  \vspace{-.2in} % corrects bad spacing
  \caption{#2.\label{fig:#3}}
  \end{figure}
}

\newcommand{\bvec}[1]{\mathbf{#1}}
\newcommand{\bhat}[1]{\mathbf{\hat{#1}}}

\begin{document}

\graphicspath{ {figures/} }

\title[Optimal Paths in R2]{Optimal Particle Paths Around Points in $\R^2$ with Constrained Acceleration}

\author{Jonathan Allen, John Wang}
\date{\today}

\maketitle

\subsection*{Abstract}

Optimal particle paths in $\R^2$ between points and around points are investigated in this paper. A paramerization of a particle's motion around a point is defined. From this parameterization, we derive a general solution for an optimal path with constant speed between 2 points. The results of this paper can by applied towards calculating optimal object trajectories in physics when acceleration is constrained.

\section{Introduction}
\input{introduction.tex}

\section{Notation}
\input{notation.tex}

\section{Particles and Paths}
\input{particle_definitions.tex}

\section{Traveling Between Points}
\input{path_theorems.tex}

\section{Turning Around Cones}
\input{cone_theorems.tex}

\section{Conclusion}
\input{conclusion.tex}

\section*{Appendix}
\input{math_review.tex}

\input{bibliography.tex}

\end{document}