### 2013-10-07 particle_definitions.tex
Finish modifying particle and path defs.

In this section, we will provide some basic definitions about particles and paths. These will lay the groundwork for thinking about optimal paths.

\begin{definition}
  A n-dimensional path $\gamma(t): \mathrm{R} \to \mathrm{R}^n$ is a function which maps a time $t \in \R$, $t \in [T_0(\gamma), T_f(\gamma)]$, to a position $\bvec{X} \in \mathrm{R}^n$. 
\end{definition}

\begin{definition}
  A planar path is a path in $\R^2$.
\end{definition}

\begin{definition}
  A path between two points, $\bvec{X_1}$ and $\bvec{X_2}$ is a path, $\gamma(t)$, where $\gamma(T_0(\gamma)) = \bvec{X_1}$ and $\gamma(T_f(\gamma)) = \bvec{X_2}$.
\end{definition}

\begin{definition}
  A particle, $p$, is an object in space with zero volume. A particle travels along a path, $\gamma(t)$ if the particle is at position $\gamma(t)$ at time $t$, for all $t \in [T_0(\gamma), T_f(\gamma)]$. If a particle is traveling along a path, $\gamma(t)$, then we define a position, $\bvec{X}(t) \coloneqq \gamma(t)$, a velocity, $\bvec{v}(t) \coloneqq \frac{d\gamma(t)}{dt}$, and an acceleration, $\bvec{a}(t) \coloneqq \frac{d^2\gamma(t)}{dt^2}$, for the particle, for all $t \in [T_0(\gamma), T_f(\gamma)]$.

  A particle is restricted if there are conditions on its position and the time derivatives of its position. Given a particle, a valid path is a path that the particle can travel along.
\end{definition}

\begin{definition}
  Given a particle, a fastest path, $\hat{\gamma}(t)$, between two points, $\bvec{X_1}$ and $\bvec{X_2}$, is a valid path such that $T_f(\hat{\gamma}) \leq T_f(\gamma)$ for all valid paths, $\gamma(t)$, between $\bvec{X_1}$ and $\bvec{X_2}$.
\end{definition}

\begin{definition}
  A particle's speed is $v(t)$, and its direction of motion is $\bhat{v}$.
\end{definition}

\begin{definition}
  The centripetal acceleration, $\bvec{a_c}$, of a particle, $p$, is the component of its acceleration in the direction perpendicular to its direction of motion.

  In rectangular coordinates, the sign of $a_c$ is defined to be the sign of the projection of $\bhat{a_c}$ onto $\bhat{x}$.

  In polar coordinates, the sign of $a_c$ is defined to be the sign of the projection of $\bhat{a_c}$ onto $\bhat{r}$.
\end{definition}

\begin{definition}
  The tangential acceleration, $a_t$, of a particle, is the component of the acceleration of the particle in its direction of motion.

  \[a_t \coloneqq \frac{dv}{dt}\]
\end{definition}

% -----------------------------------------------------------------------------

\subsection{Particle Motion in Polar Coordinates}

Unless otherwise specified, the motion of particle moving along planar paths  will be described in polar coordinates, along with an extra parameter $\theta$, as is shown in Figure \ref{fig:polar-param}.

\includefig{polar_param.eps}{A particle moving in 2-dimensional polar coordinate system centered at B}{polar-param}

$\theta(t), \phi(t) \in [-\pi, \pi]$. From now on, $\abs{\theta(t)}$ and $\abs{\phi(t)}$ will be used, since there is a symmetry in the system about $\bhat{r}$.

\begin{lemma}
The time derivative of $\theta$ is given by
\[
\frac{d\abs{\theta(t)}}{dt} = \frac{a_c}{v}
\]
\end{lemma}

\begin{proof}

If we look at a point, $p$, subject to only centripetal acceleration, $a_c$, the change in $\bvec{v}$ over an infinitesimal time, $dt$, is shown in Figure \ref{fig:theta-deriv} (the two vectors, $\bvec{v}$ and $\bvec{v}+\bvec{dv}$, are superimposed).

\includefig{theta_deriv.eps}{Tangential acceleration}{theta-deriv}

From the definition of $a_c$

\[
\frac{dv}{dt} = a_c
\]

Since $dv = d\abs{\theta} v$, then
\[
\frac{d\abs{\theta}}{dt} = \frac{a_c}{v}
\]
\end{proof}

The proof of the lemma in the case where $a_t$ is also nonzero is very similar, since the component of $\bvec{dv}$ in the direction $\bhat{v}$ is negligible compared to $v$.

Returning again to Figure \ref{fig:polar-param}, the following equations can be derived

\begin{align}
  \frac{dr(t)}{dt}& = -v(t) \, \cos\paren{\abs{\theta(t)}}\label{eq:r-deriv}\\
  \frac{d\abs{\phi(t)}}{dt}& = \frac{v(t)}{r(t)} \, \sin\paren{\abs{\theta(t)}}\label{eq:phi-deriv}\\
  \frac{d\abs{\theta(t)}}{dt}& = \frac{d\abs{\phi(t)}}{dt} - \frac{a_c(t)}{v(t)}\\
  &= \frac{v(t)}{r(t)} \, \sin\paren{\abs{\theta(t)}} - \frac{a_c(t)}{v(t)}\label{eq:theta-deriv}\\
  \intertext{Applying the chain rule to (\ref{eq:r-deriv})}
  \frac{d}{dt} \frac{dr(t)}{dt}& = -\frac{dv(t)}{dt} \cos\paren{\abs{\theta(t)}} + v(t) \sin\paren{\abs{\theta(t)}} \frac{d\abs{\theta(t)}}{dt}\\
  &\begin{aligned}
    = -a_t \cos\paren{\abs{\theta(t)}} &+ \frac{v(t)^2}{r(t)} \sin^2\paren{\abs{\theta(t)}}\\
    &\qquad - v(t) \sin\paren{\abs{\theta(t)}} \frac{a_c(t)}{v(t)}\label{eq:r-deriv-2}
  \end{aligned}
\end{align}

\begin{lemma}\label{lemma:polar-assymptote}
  Given a particle with the following restrictions: $a = 0$, $v(t) = \bar{v}$, for all $t \ge T_0$then the following equations hold
  \[
    \begin{cases}
      \abs{\theta(t)} \to \pi \quad \text{as} \quad t \to \infty \qquad &\text{if} \;\; \theta(T_0) > 0\\
      \abs{\theta(t)} = 0 \quad \text{for all} \quad t \ge T_0 \qquad &\text{if} \;\; \theta(T_0) = 0
    \end{cases}
  \]

    Furthermore,

  \[
    \begin{cases}
      r(t) \to \infty \quad \text{as} \quad t \to \infty \qquad &\text{if} \;\; \theta(T_0) > 0\\
      r(t) \to 0 \quad \text{as} \quad t \ge T_0 \qquad &\text{if} \;\; \theta(T_0) = 0
    \end{cases}
  \]
\end{lemma}

\begin{proof}

The first part of the lemma is proven by referring to (\ref{eq:theta-deriv}), and noting that $\frac{d\abs{\theta(t)}}{dt} = 0$ when $\abs{\theta(t)} = 0, \pi$, and $\frac{d\abs{\theta(t)}}{dt} > 0$ otherwise. If $\theta(T_0) \ne 0$, then $\theta(t)$ is bounded above by $\pi$ and strictly increasing, because its derivative is strictly positive, so $\abs{\theta(t)} \to \pi \quad \text{as} \quad t \to \infty$. If $\theta(T_0) = 0$, then $\abs{\theta(t)} = 0 \quad \text{for all} \quad t \ge T_0$.

For the first case of the second part of the lemma,

\begin{align*}
\frac{d\abs{\theta(t)}}{dt} = 0 \qquad \text{for} \;\; t \ge t_0\\
\intertext{So}
\abs{\theta(t)} = 0 \qquad \text{for} \;\; t \ge t_0
\end{align*}

[TODO]

\end{proof}