### 2013-10-08 math_review.tex
Remove unneeded sections of appendix.

\subsection{Coordinate Systems}

\subsection{Vector Calculus in Polar Coordinates}

\begin{align*}
\boldsymbol{x}& = r \boldsymbol{\hat{r}}\\
\boldsymbol{\vec{v}}& = \dot{r} \boldsymbol{\hat{r}} + r \dot{\phi} \boldsymbol{\hat{\phi}}\\
\label{eq:a_vec_def}\boldsymbol{\vec{a}}& = \paren{\ddot{r} - r \dot{\phi}^2} \boldsymbol{\hat{r}} + \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}} \boldsymbol{\hat{\phi}}
\end{align*}