### 2013-10-09 conclusion.tex
Fixing typos

In this paper, we have formalized many of the things that intuition would tell us. Namely, we have shown that a straight line is the fastest way to get between two points. This fact is unsurprising because of the fact that acceleration in a single direction (toward the finish point) is the least wasteful means of getting to the finish.

We developed a number of simple lemmas governing the motion of a volume-less particle with bounded acceleration. We also showed that when a particle has bounded accleration, the optimal strategy for getting to another point as fast as possible is to minimize the angle that the particle's velocity vector makes with the ending point.