### 2013-10-10 paper.tex

\documentclass[12pt]{amsart}   % LaTeX with AMS style; 12 point for old eyes

\usepackage{amsmath,amssymb,amsfonts}   % For better support of math
\usepackage{graphicx}	        % Enable for eps and pdf figures, if they occur
\usepackage{caption}
\usepackage{float}
\usepackage{subcaption}
\usepackage{epstopdf}
\usepackage{hyperref} % Enable embedded hyperlinks.
\hypersetup{
	hidelinks, colorlinks, linkcolor=black, citecolor=black, urlcolor=red
}

% Commands to force sequential numbering:

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{examples}[theorem]{Examples}
\newtheorem{remarks}[theorem]{Remarks}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}

% Define abs norm, paren, bracket, cbracket, and innerproduct
\usepackage{mathtools}
\DeclarePairedDelimiter\tempabs{\lvert}{\rvert}
\DeclarePairedDelimiter\tempnorm{\lVert}{\rVert}
\DeclarePairedDelimiter\tempinnerproduct{\langle}{\rangle}
\DeclarePairedDelimiter\tempparen{(}{)}
\DeclarePairedDelimiter\tempbracket{[}{]}
\DeclarePairedDelimiter\tempcbracket{\{}{\}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
% Swap * functionality
\makeatletter
\def\abs{\@ifstar{\tempabs}{\tempabs*}}
\def\norm{\@ifstar{\tempnorm}{\tempnorm*}}
\def\innerproduct{\@ifstar{\tempinnerproduct}{\tempinnerproduct*}}
\def\paren{\@ifstar{\tempparen}{\tempparen*}}
\def\bracket{\@ifstar{\tempbracket}{\tempbracket*}}
\def\cbracket{\@ifstar{\tempcbracket}{\tempcbracket*}}
\makeatother

% blackboard bold , for ``complex,'' etc
\newcommand{\C}{\mathbb C} 
\newcommand{\R}{\mathbb R} 
\newcommand{\Z}{\mathbb Z}
\newcommand{\Q}{\mathbb Q}
\newcommand{\N}{\mathbb N}

% Add fig with caption command. Inputs: {path}{caption}{label}
\newcommand{\includefig}[3]{
  \begin{figure}[H]
    \begin{center}
      \includegraphics[width=4in,height=3in,keepaspectratio]{#1}
    \end{center}
  \vspace{-.2in} % corrects bad spacing
  \caption{#2.\label{fig:#3}}
  \end{figure}
}

\newcommand{\bvec}[1]{\mathbf{#1}}
\newcommand{\bhat}[1]{\mathbf{\hat{#1}}}

\begin{document}

\graphicspath{ {figures/} }

\title[Fastest Paths in R2]{Fastest Paths Between Points in $\R^2$ for a Particle Moving with Constant Speed and Bounded Centripetal Acceleration}

\author{Jonathan Allen, John Wang}
\date{\today}

\maketitle

\subsection*{Abstract}

This paper investigates fastest paths for a particle moving between two points in $\R^2$. A modified polar coordinate system is defined, which we use to derive differential equations to model the particle's motion. These equations allow us to describe theorems about fastest paths between points. The results of this paper can by applied to modeling a particle's motion in $\R^2$ and to calculating shortest paths in $\R^2$.

\section{Introduction}
\input{introduction.tex}

\section{Notation}
\input{notation.tex}

\section{Particles and Paths}
\input{particle_definitions.tex}

\section{Traveling Between Points}
\input{path_theorems.tex}

\section{Conclusion}
\input{conclusion.tex}

\section*{Appendix}
\input{math_review.tex}

\input{bibliography.tex}

\end{document}