### 2013-10-11 math_review.tex

\subsection{Vectors}

\begin{description}
  \item[$\bvec{a}$] Vector in n-dimensional space.
  \item[$\bhat{a}$] Unit vector in n-dimensional space. $\bhat{a} = \bvec{a}/a$ and $\norm{\bhat{a}} = 1$.
  \item[$a$] Scalar quantity. Usually $a = \norm{\bvec{a}}$, but occasionally we will also define a sign that is dependent on the direction of $\bhat{a}$.
\end{description}

\subsection{Vector Calculus in Polar Coordinates}

\begin{align*}
\boldsymbol{x}& = r \boldsymbol{\hat{r}}\\
\boldsymbol{\vec{v}}& = \dot{r} \boldsymbol{\hat{r}} + r \dot{\phi} \boldsymbol{\hat{\phi}}\\
\label{eq:a_vec_def}\boldsymbol{\vec{a}}& = \paren{\ddot{r} - r \dot{\phi}^2} \boldsymbol{\hat{r}} + \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}} \boldsymbol{\hat{\phi}}
\end{align*}