### 2013-10-11 particle_definitions.tex
Added paper.pdf.

We will first provide some basic definitions about particles and paths, which will lay the groundwork for thinking about fastest paths. First we will define what we mean by a particle and a path.

\begin{definition}
  A path $\gamma(t): [T_0(\gamma), T_f(\gamma)] \to \mathrm{R}^2$ is a function which maps a time $t \in [T_0(\gamma), T_f(\gamma)]$ to a position $\bvec{X} \in \mathrm{R}^2$.
\end{definition}

\begin{definition}
  A particle $p$ is an object in space with zero volume. If a particle is traveling along a path, $\gamma(t)$, then we define a position, $\bvec{X}(t) \coloneqq \gamma(t)$, a velocity, $\bvec{v}(t) \coloneqq \frac{d\gamma(t)}{dt}$, and an acceleration, $\bvec{a}(t) \coloneqq \frac{d^2\gamma(t)}{dt^2}$, for the particle, for all $t \in [T_0(\gamma), T_f(\gamma)]$.

  A particle is restricted if there are conditions on its position and the time derivatives of its position. Given a particle, a valid path is a path that the particle can travel along without violating its restrictions.
\end{definition}

Now we will make our problem statement more rigorous, by defining what a fastest path is.

\begin{definition}
  Let $p$ be a particle with initial position $X_0$, a fastest path $\hat{\gamma}(t)$ to $\bvec{X_f}$, is a valid path such that $T_f(\hat{\gamma}) \leq T_f(\gamma)$ for all valid paths $\gamma(t)$ to $\bvec{X_f}$.
\end{definition}

Finally, we need to define some terms for a particle's motion which will be needed later on to describe restrictions on a particle's motion.

\begin{definition}
  The centripetal acceleration $\bvec{a_c}(t)$ of a particle, $p$, is the component of its acceleration in the direction perpendicular to its direction of motion.
\end{definition}

\begin{definition}
  The tangential acceleration $a_t(t)$ of a particle is the component of the acceleration of the particle in its direction of motion.

  \begin{equation}
    a_t(t) \coloneqq \frac{dv(t)}{dt}
  \end{equation}
\end{definition}
% -----------------------------------------------------------------------------

\subsection{Particle Motion in Polar Coordinates}

Before we introduce the equations governing a particle's motion in $\R^2$, we must first define a coordinate system to describe its motion. Unless otherwise specified, the motion of particles in this paper will be described in a modified polar coordinate system that contain an extra parameter $\theta$, that is shown in Figure \ref{fig:polar-param}. We will refer to this system as the \textit{B centered particle coordinate system}.

\begin{figure}[H]
    \begin{center}
      \includegraphics[width=4in,height=3in,keepaspectratio]{polar_param.eps}
    \end{center}
  \vspace{-.2in} % corrects bad spacing
  \caption{A particle moving in a B centered particle coordinate system.\label{fig:polar-param}}
  \end{figure}

The sign of $a_c(t)$ is defined to be the sign of the projection of $\bhat{a_c}(t)$ onto $-\bhat{r}(t)$; in other words, it is positive when pointing towards B. $\theta(t), \phi(t) \in [-\pi, \pi]$, and from now on, we will only use $\abs{\theta(t)}$ and $\abs{\phi(t)}$, since there is a symmetry in the system about $\bhat{r}$.

\begin{lemma}
The time derivative of $\abs{\theta(t)}$ is given by

\begin{equation}\label{eq:theta-deriv}
\frac{d\abs{\theta(t)}}{dt} = \frac{d\abs{\phi(t)}}{dt} + \frac{a_c(t)}{v(t)}
\end{equation}

\end{lemma}

\begin{proof}

$\theta(t)$ is composed of $\phi(t)$ plus the angle between $\bhat{v}(t)$ and the horizontal, which we can temporarily call $\beta(t)$. $\phi(t)$ is the component of $\theta(t)$ in the polar coordinate system, and changes as the point moves with respect to the point B. $\beta(t)$ can be thought of as the component of $\theta(t)$ in the Cartesian coordinate system, and it only changes when $a_c(t)$ is nonzero.

Differentiating, we get

\[
\frac{d\abs{\theta(t)}}{dt} = \frac{d\abs{\phi(t)}}{dt} + \frac{d\beta(t)}{dt}
\]

To find $\frac{d\beta(t)}{dt}$, we can look at a point, $p$, subject to nonzero centripetal acceleration, and zero tangential acceleration. The change in $\bvec{v(t)}$ over an infinitesimal time, $dt$, is shown below (the two vectors, $\bvec{v(t)}$ and $\paren{\bvec{v(t)}+\bvec{dv(t)}}$, are superimposed).

\begin{figure}[H]
    \begin{center}
      \includegraphics[width=3.5in,height=2.5in,keepaspectratio]{theta_deriv.eps}
    \end{center}
  \vspace{-.2in} % corrects bad spacing
  \caption{}
\end{figure}

$a_c(t)$ is defined to be the component of acceleration perpendicular to the direction of motion, so

\[
\frac{dv(t)}{dt} = a_c(t)
\]

Since $dv(t) = -d\beta(t) v(t)$, then

\[
\frac{d\beta(t)}{dt} = -\frac{a_c(t)}{v(t)}
\]

The proof of the lemma in the case where $a_t(t)$ is nonzero is nearly identical, since the component of $\bvec{dv}(t)$ in the direction $\bhat{v}(t)$ is negligible compared to $v(t)$.

\end{proof}

Returning again to Figure \ref{fig:polar-param}, the following equations can be derived

\begin{align}
  \frac{dr(t)}{dt}& = -v(t) \, \cos\paren{\abs{\theta(t)}}\label{eq:r-diff}\\
  \frac{d\abs{\phi(t)}}{dt}& = \frac{v(t)}{r(t)} \, \sin\paren{\abs{\theta(t)}}\label{eq:phi-diff}\\
  \frac{d\abs{\theta(t)}}{dt}& = \frac{d\abs{\phi(t)}}{dt} - \frac{a_c(t)}{v(t)}\\
  &= \frac{v(t)}{r(t)} \, \sin\paren{\abs{\theta(t)}} - \frac{a_c(t)}{v(t)}\label{eq:theta-diff}\\
  \intertext{Applying the chain rule to (\ref{eq:r-diff})}
  \frac{d}{dt} \frac{dr(t)}{dt}& = -\frac{dv(t)}{dt} \cos\paren{\abs{\theta(t)}} + v(t) \sin\paren{\abs{\theta(t)}} \frac{d\abs{\theta(t)}}{dt}\\
  &\begin{aligned}
    = -a_t(t) \cos\paren{\abs{\theta(t)}} &+ \frac{v(t)^2}{r(t)} \sin^2\paren{\abs{\theta(t)}}\\
    &\qquad - \sin\paren{\abs{\theta(t)}} a_c(t) \label{eq:r-diff-2}
  \end{aligned}
\end{align}